---
title: "Portfolio 3"
---

> This project aims to create reusable codes that create the inter-response time plot for the DRL-20 paradigm for my first year projects. My first year project looks at the effect of acetylcholine antagonists in the nucleus accumbens on rats' impulsive action, measured by the DRL-20 paradigm (restrain lever press for 20 seconds to get a sugar pallet). The 3 main DVs we're looking at is reinforcerns earned, lever presses, and efficiency (reinforcers earned/total level presses), which is easy to calculate through Excel or SPSS. The inter-reponse time plot, however, require some data wrangling and is a pain in the ass to do in Excel, so that's why I want to use R to make it easier here. 

```{r load-in-packages}
library(readxl)
library(tidyverse)
```

```{r read-in-data}
IE_CArray_DD1 <- read_excel("data/IE_CArray_DD1.xlsx")

#clean the rows
table(IE_CArray_DD1$Subject) 
#11 rows that is the header, need to filter them out
IE_D1 <- IE_CArray_DD1 %>%
  filter(!(Subject %in% "Subject"))
as_tibble(IE_D1)
#check the data
table(IE_D1$Subject) #looks good, now I need to convey Box and Carray to numbers
IE_D1$Box <- as.numeric(IE_D1$Box)
IE_D1$Carray <- as.numeric(IE_D1$Carray)
```

```{r clean-the-data}
IE_D1$time <- round(IE_D1$Carray)/100 
IE_D1$interval <-  c(0, diff(IE_D1$time))
IE_D1$interval <- ifelse(IE_D1$interval < 0, 0, IE_D1$interval)
```

```{r check}
table(IE_D1$Subject, IE_D1$Drug)
#IE_D1$divisor <- ifelse(IE)
```
From the table, we can tell that 1 rat got High dose on that day, 4 got Low dose, 3 got Medium dose, and 4 got Vehicle (saline control).

```{r plot-histogram}
hist <- ggplot(IE_D1,
    aes(x = interval, fill = Drug)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ Drug, nrow = 4) +
  theme_minimal()
hist

#get bin counts
bin_data <- ggplot_build(hist)$data[[1]]
head(bin_data)

#create count1 as the average interval across rats
bin_data$count1 <- ifelse(bin_data$PANEL == 2, bin_data$count/4, bin_data$count)
bin_data$count1 <- ifelse(bin_data$PANEL == 3, bin_data$count1/3, bin_data$count1)
bin_data$count1 <- ifelse(bin_data$PANEL == 4, bin_data$count1/4, bin_data$count1)
bin_data$count1 <- round(bin_data$count1)

#create drug variable
bin_data$drug <- ifelse(bin_data$PANEL == 1, "High dose", 0)
bin_data$drug <- ifelse(bin_data$PANEL == 2, "Lose dose", bin_data$drug)
bin_data$drug <- ifelse(bin_data$PANEL == 3, "Medium dose", bin_data$drug)
bin_data$drug <- ifelse(bin_data$PANEL == 4, "Vehicle", bin_data$drug)

ggplot(bin_data, aes(x = x, y = count1, fill = drug)) +
  geom_col() +
  facet_wrap(~ group, nrow = 4) +
  labs(title = "Inter-Response Time for Four Doses of Drugs",
       x = "Inter-Response Time",
       y = "Count", 
       fill = "Drug") +
  theme_minimal()
```

```{r plot-density}
ggplot(IE_D1,
    aes(x = interval, color = Drug)) +
  geom_density(linewidth = 1, alpha = .5) +
  theme_minimal() 
#+ scale_color_viridis_d() #for discrete color

ggplot(bin_data,
    aes(x = x, y = count1, color = factor(group))) +
  geom_smooth(method = "loess") +
  theme_minimal()
```

```{r plot-histogram-density}
#ggplot(IE_D1,
#    aes(x = interval, fill = Drug)) +
#  geom_histogram(binwidth = 1, alpha = .5, position = "identity") +
#  geom_line(aes(y = ..count.., linetype = Drug)) +
#  theme_minimal() +
#  labs(title = "Histogram with Scaled Density Line", x = "Values", y = "Count/Density")
#+ scale_color_viridis_d()

ggplot(bin_data, aes(x = x, y = count1, linetype = drug, color = drug)) +
  geom_line(size = 1) +
  labs(title = "Inter-Response Time for Four Doses of Drugs",
       x = "Inter-Response Time",
       y = "Count", 
       fill = "Drug") +
  theme_minimal()
```

