---
title: "Portfolio 4"
---

> This project aims to analyze the main DVs of my first year project, which looks at the effect of acetylcholine antagonists in the nucleus accumbens on rats' impulsive action, measured by the DRL-20 paradigm (restrain lever press for 20 seconds to get a sugar pallet). The 3 main DVs we're looking at is reinforcerns earned, lever presses, and efficiency (reinforcers earned/total level presses)

### 0. Prep & data cleaning.

```{r load-in-packages}
library(readxl)
library(tidyverse)
library(haven)
library(lme4)
library(lmerTest)
library(ez)
library(emmeans)
library(effectsize)
```

```{r read-in-data}
IM_DD1_BasicData <- read_excel("data/fyp/IM_DD1_BasicData.xlsx")
IM_DD2_BasicData <- read_excel("data/fyp/IM_DD2_BasicData.xlsx")
IM_DD3_BasicData <- read_excel("data/fyp/IM_DD3_BasicData.xlsx")
IM_DD4_BasicData <- read_excel("data/fyp/IM_DD4_BasicData.xlsx")
IM_DD5_BasicData <- read_excel("data/fyp/IM_DD5_BasicData.xlsx")
```

```{r combine-the-data}
IM_all <- bind_rows(IM_DD1_BasicData, IM_DD2_BasicData, IM_DD3_BasicData, IM_DD4_BasicData, IM_DD5_BasicData)
```

```{r clean-data}
IM_all$Group <- ifelse(IM_all$Group %in% c("M_HI", "MEC_HI", "MEC-HI"), "M_HIGH", IM_all$Group)
IM_all$Group <- ifelse(IM_all$Group %in% c("M_LO", "MEC_LO", "MEC-LO", "MEC-LOW"), "M_LOW", IM_all$Group)
IM_all$Group <- ifelse(IM_all$Group %in% c("S_HI", "SCOP_HI", "SCOP-HI"), "S_HIGH", IM_all$Group)
IM_all$Group <- ifelse(IM_all$Group %in% c("S_LO", "SCOP_LO", "SCOP-LOW", "SCP-LOW", "SCOP-LO"), "S_LOW", IM_all$Group)
```

```{r cal-efficiency}
IM_all$Efficiency <- IM_all$ReinforcersEarned / IM_all$RightLeverPress * 100
```

```{r write-out-data}
#long to wide
IM_all_wide <- IM_all %>%
  select(Subject, Group, Efficiency, ReinforcersEarned, RightLeverPress, LeftLeverPress, NosePokes) %>%
  pivot_wider(
    names_from = Group,  # Columns to create
    values_from = c(Efficiency, ReinforcersEarned, RightLeverPress, LeftLeverPress, NosePokes)     # Values to fill
  )

write_sav(IM_all_wide, "data/fyp/IM_all_wide.sav")
```

```{r exclude-IM10}
IM_all_clean <- IM_all %>%
  filter(Subject != "IM10")
```


### 1. Visualization.
```{r vis}
IM_all_clean$Group <- factor(IM_all_clean$Group, levels = c("M_HIGH", "M_LOW", "VEH", "S_LOW", "S_HIGH"))
write_rds(IM_all_clean, "data/fyp/IM_all_clean.rds")

#Lever press plot
IM_all_clean %>%
  ggplot(aes(x = Group, y = RightLeverPress)) +
  geom_boxplot() +
  geom_point(aes(color = Subject)) +  
  stat_summary(fun = mean, geom = "line", group = 1, color = "black") +  
  stat_summary(fun = mean, geom = "point", color = "black") 

#Reinforcers earned plot
IM_all_clean %>%
  ggplot(aes(x = Group, y = ReinforcersEarned)) +
  geom_boxplot() +
  geom_point(aes(color = Subject)) +  
  stat_summary(fun = mean, geom = "line", group = 1, color = "black") +  
  stat_summary(fun = mean, geom = "point", color = "black") 

#Efficiency plot
IM_all_clean %>%
  ggplot(aes(x = Group, y = Efficiency)) +
  geom_boxplot() +
  geom_point(aes(color = Subject)) +  
  stat_summary(fun = mean, geom = "line", group = 1, color = "black") +  
  stat_summary(fun = mean, geom = "point", color = "black")  
```

```{r multilevel-model, eval=FALSE}
#random intercept
LPmodel <- lmer(RightLeverPress ~ Group + (1|Subject), data = IM_all_clean)
summary(LPmodel)

#random slope
LPmodel.random.slope <- lmer(RightLeverPress ~ Group + (Group|Subject), data = IM_all_clean)
summary(LPmodel.random.slope)

#likelihood ratio test
anova(LPmodel.random.slope, LPmodel)
```
Multilevel modelling doesn't work because we don't have enough data points. 

### ANOVA and post-hocs.
```{r anova-t-test}
#anova: Right Lever Press
RLP_results <- ezANOVA(
  data = IM_all_clean,
  dv = .(RightLeverPress),     # Dependent variable
  wid = .(Subject),  # Within-subject variable (ID)
  within = .(Group)   # Repeated measures factor
)
RLP_results #sig
#post hoc
pairwise.t.test(IM_all_clean$RightLeverPress, IM_all_clean$Group, paired = TRUE, p.adjust.method = "bonferroni")
#S_HIGH pressed significantly less levers than all other groups
###################################################################################
# Conduct repeated measures ANOVA
anova_result <- aov_car(Efficiency ~ Group + Error(Subject/Group), data = IM_all_clean)

# View ANOVA table
print(anova_result)

# Calculate effect size
effect_sizes <- eta_squared(anova_result)
print(effect_sizes)
###################################################################################

#anova: Reinforcers Earned
RE_results <- ezANOVA(
  data = IM_all_clean,
  dv = .(ReinforcersEarned),     # Dependent variable
  wid = .(Subject),  # Within-subject variable (ID)
  within = .(Group)   # Repeated measures factor
)
RE_results #sig
#post hoc
pairwise.t.test(IM_all_clean$ReinforcersEarned, IM_all_clean$Group, paired = TRUE, p.adjust.method = "bonferroni")

# Fit the repeated measures model
RE_model <- aov(ReinforcersEarned ~ Group + Error(Subject / Group), data = IM_all_clean)

# Pairwise comparisons with Tukey adjustment
RE_tukey <- emmeans(RE_model, pairwise ~ Group, adjust = "tukey")

# View results
RE_tukey

#S_HIGH got significantly less sugar pallets than all other groups

#anova: Efficiency
E_results <- ezANOVA(
  data = IM_all_clean,
  dv = .(Efficiency),     # Dependent variable
  wid = .(Subject),  # Within-subject variable (ID)
  within = .(Group)   # Repeated measures factor
)
E_results #not sig
#post hoc
pairwise.t.test(IM_all_clean$Efficiency, IM_all_clean$Group, paired = TRUE, p.adjust.method = "bonferroni")
#all groups have the same efficiency

#anova: Nose pokes
NP_results <- ezANOVA(
  data = IM_all_clean,
  dv = .(NosePokes),     # Dependent variable
  wid = .(Subject),  # Within-subject variable (ID)
  within = .(Group)   # Repeated measures factor
)
NP_results #sig
#post hoc
pairwise.t.test(IM_all_clean$NosePokes, IM_all_clean$Group, paired = TRUE, p.adjust.method = "bonferroni")
#rats who received scopolamine high had less nose pokes
```

```{r two-drug-sep}
#if efficiency for two drugs are analyzed separately 
#sco
E_sco_results <- IM_all_clean %>%
  filter(Group %in% c("VEH", "S_LOW", "S_HIGH")) %>%
  ezANOVA(
  dv = .(Efficiency),     # Dependent variable
  wid = .(Subject),  # Within-subject variable (ID)
  within = .(Group)   # Repeated measures factor
)
E_sco_results #not sig

#mec
E_mec_results <- IM_all_clean %>%
  filter(Group %in% c("VEH", "M_LOW", "M_HIGH")) %>%
  ezANOVA(
  dv = .(Efficiency),     # Dependent variable
  wid = .(Subject),  # Within-subject variable (ID)
  within = .(Group)   # Repeated measures factor
)
E_mec_results #not sig
```

### Conclusion.
1. Drug condition significantly affects lever presses. Specifically, high dose of scopolamine decreases lever presses.
2. Drug condition significantly affects reinforcers earned. Specifically, rats who got high dose of scopolamine received less sugar pallets.
3. Drug condition doesn't affect efficiency.

Together, the results implied that high dose of scopolamine might impair rats' motor activities. Further analyses are needed.

### Now some bar plots.

```{r bar-plot-lever-press}
# total level press
# Calculate standard error
means <- IM_all_clean %>%
  group_by(Group) %>%
  summarize(
    MeanScore = mean(RightLeverPress),
    SE = sd(RightLeverPress) / sqrt(n()),
    .groups = "drop"
  )

# Add error bars
ggplot(means, aes(x = Group, y = MeanScore, fill = Group)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = MeanScore - SE, ymax = MeanScore + SE), width = 0.15, color = "darkgrey") +
  labs(title = "Total Lever Presses", x = NULL, y = NULL) +
  scale_fill_manual(values = c("M_HIGH" = "#397DB0", "M_LOW" = "#9DC7DD", "VEH" = "#fbf0d9", "S_LOW" = "#9ED17B", "S_HIGH" = "#3D9F3C")) +
  scale_x_discrete(labels = c("M_HIGH" = "Mecamylamine\nHigh Dose", "M_LOW" = "Mecamylamine\nLow Dose", "VEH" = "Vehicle", "S_LOW" = "Scopolamine\nLow Dose", "S_HIGH" = "Scopolamine\nHigh Dose")) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 11, face = "bold"), axis.text.y = element_text(size = 15, face = "bold"),  plot.title = element_text(hjust = 0.5, size = 20, face = "bold"))
```

```{r bar-plot-reinforcers-earned}
# total reinforcers earned
# Calculate standard error
means <- IM_all_clean %>%
  group_by(Group) %>%
  summarize(
    MeanScore = mean(ReinforcersEarned),
    SE = sd(ReinforcersEarned) / sqrt(n()),
    .groups = "drop"
  )

# Add error bars
ggplot(means, aes(x = Group, y = MeanScore, fill = Group)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = MeanScore - SE, ymax = MeanScore + SE), width = 0.15, color = "darkgrey") +
  labs(title = "Reinforcers Earned", x = NULL, y = NULL) +
  scale_fill_manual(values = c("M_HIGH" = "#397DB0", "M_LOW" = "#9DC7DD", "VEH" = "#fbf0d9", "S_LOW" = "#9ED17B", "S_HIGH" = "#3D9F3C")) +
  scale_x_discrete(labels = c("M_HIGH" = "Mecamylamine\nHigh Dose", "M_LOW" = "Mecamylamine\nLow Dose", "VEH" = "Vehicle", "S_LOW" = "Scopolamine\nLow Dose", "S_HIGH" = "Scopolamine\nHigh Dose")) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 11, face = "bold"), axis.text.y = element_text(size = 15, face = "bold"),  plot.title = element_text(hjust = 0.5, size = 20, face = "bold")) 
```

```{r bar-plot-efficiency}
# efficiency
# Calculate standard error
means <- IM_all_clean %>%
  group_by(Group) %>%
  summarize(
    MeanScore = mean(Efficiency),
    SE = sd(Efficiency) / sqrt(n()),
    .groups = "drop"
  )

# Add error bars
ggplot(means, aes(x = Group, y = MeanScore, fill = Group)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = MeanScore - SE, ymax = MeanScore + SE), width = 0.15, color = "darkgrey") +
  labs(title = "Efficiency", x = NULL, y = NULL) +
  scale_fill_manual(values = c("M_HIGH" = "#397DB0", "M_LOW" = "#9DC7DD", "VEH" = "#fbf0d9", "S_LOW" = "#9ED17B", "S_HIGH" = "#3D9F3C")) +
  scale_x_discrete(labels = c("M_HIGH" = "Mecamylamine\nHigh Dose", "M_LOW" = "Mecamylamine\nLow Dose", "VEH" = "Vehicle", "S_LOW" = "Scopolamine\nLow Dose", "S_HIGH" = "Scopolamine\nHigh Dose")) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 11, face = "bold"), axis.text.y = element_text(size = 15, face = "bold"),  plot.title = element_text(hjust = 0.5, size = 20, face = "bold")) 
```


```{r bar-plot-nose-poke}
# total nose pokes
# Calculate standard error
means <- IM_all_clean %>%
  group_by(Group) %>%
  summarize(
    MeanScore = mean(NosePokes),
    SE = sd(NosePokes) / sqrt(n()),
    .groups = "drop"
  )

# Add error bars
ggplot(means, aes(x = Group, y = MeanScore, fill = Group)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = MeanScore - SE, ymax = MeanScore + SE), width = 0.15, color = "darkgrey") +
  labs(title = "Total Nose Pokes", x = NULL, y = NULL) +
  scale_fill_manual(values = c("M_HIGH" = "#397DB0", "M_LOW" = "#9DC7DD", "VEH" = "#fbf0d9", "S_LOW" = "#9ED17B", "S_HIGH" = "#3D9F3C")) +
  scale_x_discrete(labels = c("M_HIGH" = "Mecamylamine\nHigh Dose", "M_LOW" = "Mecamylamine\nLow Dose", "VEH" = "Vehicle", "S_LOW" = "Scopolamine\nLow Dose", "S_HIGH" = "Scopolamine\nHigh Dose")) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 15),  plot.title = element_text(hjust = 0.5, size = 20, face = "bold")) 

```
