---
title: "Portfolio 7"
---

> Again, this project aims to create the right inter-response time plot for my first year project, because, obviously, I made some mistakes in portfolio piece 05 was not correct, probobly for some data management or coding problem. So this the my third attempt to create a correct IRT plot, with pre-cleaned datasets from Excel. Sometimes the simplest way might be the best way...

### Prep.

```{r read-in-data}
library(readxl)
library(tidyverse)

IM_DD1_ArrayC_clean <- read_excel("data/fyp/IRT/IM_DD1_ArrayC_clean.xlsx")
IM_DD2_ArrayC_clean <- read_excel("data/fyp/IRT/IM_DD2_ArrayC_clean.xlsx")
IM_DD3_ArrayC_clean <- read_excel("data/fyp/IRT/IM_DD3_ArrayC_clean.xlsx")
IM_DD4_ArrayC_clean <- read_excel("data/fyp/IRT/IM_DD4_ArrayC_clean.xlsx")
IM_DD5_ArrayC_clean <- read_excel("data/fyp/IRT/IM_DD5_ArrayC_clean.xlsx")
```

```{r combine}
# combine Carray data from all five days
IM_Carray_all_clean <- bind_rows(IM_DD1_ArrayC_clean, IM_DD2_ArrayC_clean, IM_DD3_ArrayC_clean, IM_DD4_ArrayC_clean, IM_DD5_ArrayC_clean)

# check the dataset to make sure there's no wierd rows
table(IM_Carray_all_clean$Subject)
IM_Carray_all_clean %>%
  count(Carray == 0)

# all looks good! Now change the column name to something more intuitive
IM_Carray_all_clean <- IM_Carray_all_clean %>%
  rename(
    Day = Experiment
  )
```

```{r clean-data}
IM_Carray_all_clean$Drug <- ifelse(IM_Carray_all_clean$Drug %in% c("M_HI", "MEC_HI", "MEC-HI"), "M_HIGH", IM_Carray_all_clean$Drug)
IM_Carray_all_clean$Drug <- ifelse(IM_Carray_all_clean$Drug %in% c("M_LO", "MEC_LO", "MEC-LO", "MEC-LOW"), "M_LOW", IM_Carray_all_clean$Drug)
IM_Carray_all_clean$Drug <- ifelse(IM_Carray_all_clean$Drug %in% c("S_HI", "SCOP_HI", "SCOP-HI"), "S_HIGH", IM_Carray_all_clean$Drug)
IM_Carray_all_clean$Drug <- ifelse(IM_Carray_all_clean$Drug %in% c("S_LO", "SCOP_LO", "SCOP-LOW", "SCP-LOW", "SCOP-LO"), "S_LOW", IM_Carray_all_clean$Drug)

# exclude IM10, who had some brain damage per histology
IM_Carray_all_clean <- IM_Carray_all_clean %>%
  filter(Subject != "IM10")

# check
table(IM_Carray_all_clean$Drug)
```

Carray is a compound variable where each event time in the box is saved with an event code:
.1 = Right lever press
.2 = Left Lever press
.3 = Nosepoke
.4 = FREE PELLET
.5 = EARNED PELLET
The number BEFORE the decimal is a time stamp. The two rightmost numbers are .XY
For example, 1800 would be 18.00 seconds.

So now ew should decompose them into different variables: Time and Event.

```{r decompose-carray}
# Separate the 'Carray' column into 'Time' and 'Event.'
IM_Carray_all_clean <- IM_Carray_all_clean %>%
  separate(Carray, into = c("Time", "Event"), sep = "\\.")

IM_Carray_all_clean$Time <- as.numeric(IM_Carray_all_clean$Time)
IM_Carray_all_clean$Time <- IM_Carray_all_clean$Time / 100

# check again, there should be no 4 in the Event row
table(IM_Carray_all_clean$Event)
# looks good!

# now export the dataset for future use
write_rds(IM_Carray_all_clean, "data/fyp/IM_Carray_all_clean.rds")

# most importantly! We need to filter out all rows that are not lever presses! Forgetting to do so before caused me to plot the wrong plot! Because I treated all events as lever presses!
IM_Carray_all_clean <- IM_Carray_all_clean %>%
  filter(Event == 1)
```

Now we need to compute the intervals for the IRT plot.
```{r compute-interval}
#calculate the inter-response time
IM_Carray_all_clean <- IM_Carray_all_clean %>%
  group_by(Subject, Day) %>%
  mutate(Interval = c(0, diff(Time))) %>%
  ungroup()

# check if there is any weird data points 
IM_Carray_all_clean %>%
  filter(Interval < 0) %>%
  count()
# looks good! no negative intervals!
IM_Carray_all_clean %>%
  filter(Interval == 0) %>%
  count()
# a lot of 0s, but I guess it makes sense that sometimes rats press levers a few times in a row

# examines intervals descriptively
interval_count <- IM_Carray_all_clean %>%
  group_by(Drug) %>%
  mutate(Interval = round(Interval)) %>%
  count(Interval)
```

Now the basic data management is done. We need to get the frequency percentage of each bin (2s) to prepare for the IRT plot.

```{r cal-bin}
# Create bins for intervals with a width of 2
IM_Carray_all_clean <- IM_Carray_all_clean %>%
  mutate(IntervalBin = floor(Interval / 2) * 2)  # Creates bins of width 2

# Calculate frequency of intervals within each bin and drug condition
bins <- IM_Carray_all_clean %>%
  group_by(Drug, IntervalBin) %>%
  summarise(Frequency = n(), .groups = "drop")

# calculate percentage
bins <- bins %>%
  group_by(Drug) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100)  # Calculate percentage
```

```{r plot}
# Create the percentage plot
bins %>%
  filter(IntervalBin <= 40) %>%
ggplot(aes(x = IntervalBin, y = Percentage, color = Drug, group = Drug)) +
  geom_line() +  # Add lines to the plot
  geom_point() +  # Optionally add points to show the frequency at each bin
  labs(
    title = "Inter-Response Times Following Acetylcholine Receptor Blockade",
    x = "Seconds",
    y = "Percentage Responses",
    color = NULL
  ) +
  scale_color_manual(
    values = c("M_HIGH" = "#397DB0", "M_LOW" = "#9DC7DD", "VEH" = "#fbf0d9", 
               "S_LOW" = "#9ED17B", "S_HIGH" = "#3D9F3C"),  # Custom colors
    labels = c("M_HIGH" = "Mecamylamine High", 
               "M_LOW" = "Mecamylamine Low", 
               "VEH" = "Vehicle", 
               "S_LOW" = "Scopolamine Low", 
               "S_HIGH" = "Scopolamine High")
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(bins$IntervalBin), max(bins$IntervalBin), by = 2)) +  # Set x-axis breaks at intervals of 2
  theme(
    legend.position = c(1, 1),  # Position the legend at the top-right corner
    legend.justification = c(1, 1),  # Align the legend to the top-right corner
    axis.text.x = element_text(size = 11, face = "bold"), 
    axis.text.y = element_text(size = 15, face = "bold"),  
    axis.title.x = element_text(face = "bold", size = 15), 
    axis.title.y = element_text(face = "bold", size = 15), 
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
  )

# Create the Frequency plot
bins %>%
  filter(IntervalBin <= 40) %>%
ggplot(aes(x = IntervalBin, y = Frequency, color = Drug, group = Drug)) +
  geom_line() +  # Add lines to the plot
  geom_point() +  # Optionally add points to show the frequency at each bin
  labs(
    title = "Inter-response Times Following Acetylcholine Receptor Blockade",
    x = NULL,
    y = "Total Response Frequency",
    color = NULL
  ) +
  scale_color_manual(
    values = c("M_HIGH" = "#397DB0", "M_LOW" = "#9DC7DD", "VEH" = "#fbf0d9", 
               "S_LOW" = "#9ED17B", "S_HIGH" = "#3D9F3C"),  # Custom colors
    labels = c("M_HIGH" = "Mecamylamine High", 
               "M_LOW" = "Mecamylamine Low", 
               "VEH" = "Vehicle", 
               "S_LOW" = "Scopolamine Low", 
               "S_HIGH" = "Scopolamine High")
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(bins$IntervalBin), max(bins$IntervalBin), by = 2)) +  # Set x-axis breaks at intervals of 2
  theme(
    legend.position = c(1, 1),  # Position the legend at the top-right corner
    legend.justification = c(1, 1),  # Align the legend to the top-right corner
    axis.text.x = element_text(size = 11, face = "bold"), 
    axis.text.y = element_text(size = 15, face = "bold"),  
    axis.title.x = element_text(face = "bold", size = 15), 
    axis.title.y = element_text(face = "bold", size = 15), 
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
  )
```

Yeah! This is what we are wanted to build!